An Improved Method for Coupling Hydrodynamics with Astrophysical Reaction Networks

### 2021-12-29 simplified_sdc_intro/paper.tex
more work

The initial model is generated following the procedure described in
\citep{subchandra}.  We use a $1.1~M_\odot$ carbon white dwarf with a
$0.05~M_\odot$ He layer.  We seed the He layer with 1\% (by mass) of
\isot{N}{14}.  The white dwarf is isothermal and the temperature ramps
up at the base of the He layer and is then isentropic until the
surface.  Figure~\ref{fig:subch_initial_model} shows the structure of
the initial model.  We use 2d axisymmetric coordinates with a domain
size of $1.024\times 10^9~\mathrm{cm} \times 2.048\times
10^9~\mathrm{cm}$, a base grid of $256 \times 512$ zones and two
levels of refinement (both jumping by $2\times$), giving a fine-grid
resolution of $10~\mathrm{km}$.  Gravity is modeled as a monopole.

@@

Overview plots of CFL = 0.8

@@

Mass of different species (as a plot)

>>>>

The initial model is generated following the procedure described in
\citep{subchandra}.  We use a $1.1~M_\odot$ carbon white dwarf with a
$0.05~M_\odot$ He layer.  We seed the He layer with 1\% (by mass) of
\isot{N}{14}.  The white dwarf is isothermal and the temperature ramps
up at the base of the He layer and is then isentropic until the
surface.  Figure~\ref{fig:subch_initial_model} shows the structure of
the initial model.  We use 2d axisymmetric coordinates with a domain
size of $1.024\times 10^9~\mathrm{cm} \times 2.048\times
10^9~\mathrm{cm}$, a base grid of $256 \times 512$ zones and two
levels of refinement (both jumping by $2\times$), giving a fine-grid
resolution of $10~\mathrm{km}$.  Gravity is modeled as a monopole.
Our choice of initial perturbation has the detonation propagate inward
into the carbon white dwarf.  We run the simulation for 0.1 s---this
is enough time for the carbon detonation to be well developed.

@@

Figure \ref{fig:subch_strang_cfl08_I} shows the temperature, mean
molecular weight, and nuclear energy generation rate for the strang
CFL = 0.8 simulation ({\tt strang\_subch2\_cfl0.8}).  We see that the
mean molecular weight appears mottled behind the inward propagating
detonation in the white dwarf.  Comparing to the energy generation
rate, we see that region is characterized by large energy releases of
alternating signs checkerboarding thoughout the region.  This is
characteristic of the nucleosynthesis struggling to come into
equilibrium at high temperatures when the reverse rates can be
important.  In Figure~\ref{fig:subch_strang_cfl08_II} we look at some
important mass fractions for the same simulation.  Again we see that
the abundances are not smooth behind the inward detonation.

\begin{figure}
\centering
\plotone{strang_subch2_cfl0.8_state}
\caption{\label{fig:subch_strang_cfl08_I}}
\end{figure}

\begin{figure}
\centering
\plotone{strang_subch2_cfl0.8_mass_fractions}
\caption{\label{fig:subch_strang_cfl08_II}}
\end{figure}

We next look at the simplified-SDC version of this same case, {\tt
  sdc\_subch2\_cfl0.8}.  Figure~\ref{fig:subch_sdc_cfl08_I} shows the
temperature, mean molecular weight, and energy generation rate.
Compared to Figure~\ref{fig:subch_strang_cfl08_I}, we see that the
composition and energy generation rates are smooth, without any of the
checkerboarding that plagued the Strang simulation.  The compositions
plot (figure~\ref{fig:subch_sdc_cfl08_II}) shows the same behavior.
This suggests that the simplified-SDC algorithm has an easier time
attaining the correct equilibrium in these conditions.

\begin{figure}
\centering
\plotone{sdc_subch2_cfl0.8_state}
\caption{\label{fig:subch_sdc_cfl08_I}}
\end{figure}

\begin{figure}
\centering
\plotone{sdc_subch2_cfl0.8_mass_fractions}
\caption{\label{fig:subch_sdc_cfl08_II}}
\end{figure}

@@

For supernova calculations, we often want to know how much mass of
\isot{Ni}{56} was produced.  Table \ref{table:masses} shows the total
mass of \isot{He}{4}, \isot{Si}{28}, and \isot{Ni}{56} for each of the
simulations.
